An Open-Source Implementation of the CMPS Algorithm for Assessing  Similarity of Bullet Signatures

### 2021-05-27 CMPSpaper_writing/CMPSpaper_writing.Rmd
added plots

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(bulletxtrctr)

@@

3. **determine the congruent registration position:** Each candidate position corresponding to each of the basis segments gets one `vote`. Votes for candidate positions within a tolerance of `Tx` are tallied and the position with the highest number of votes gets chosen as the *congruent registration position*. Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score of the comparison. \hh{XXX what happens when there are ties after the vote?}

@@

* `get_ccp(ccr.list, Tx = 25)` tries to identify the "consistent correlation peak" (defined in algorithm step 2) under the framework of multi segment lengths strategy. `ccr.list` is described in the previous bullet point, and `Tx` determines the size of a tolerance zone used in identifying the consistent correlation peak. \hh{XXX what happens when there is no consistent correlation peak?}

@@

The signatures of Land 4 of Bullet 1 and Land 5 of Bullet 2 are stored in objects sigs2 and sigs1, respectively. This comparison is one of the KM comparisons. We compute the CMPS score using two versions of the CMPS algorithm:

@@

```{r, eval=FALSE, echo=FALSE}
hamby252_results <- readRDS("C:/Research/Bullet/Bullet_Rcode/hamby252_results.rds")
hamby44_results <- readRDS("C:/Research/Bullet/Bullet_Rcode/hamby44_results.rds")
```

@@

In this example, $\mathrm{CMPS_{max}}$, the highest CMPS score is 17, and $\mathrm{\overline{CMPS}_{max}}$, the highest possible mean CMPS score of comparisons in the same phase, is $(3+17+14+10+15+16)/6=12.5$.

@@

```{r, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
# rtt <- readRDS("summary_table.rds")
# colnames(rtt) <- 
rtt <- data.frame(parameters = c("span1=0.25, (5, 3, 1)", "span1=0.25, (10, 6, 1)", "span1=0.75, (5, 3, 1)"),
                  p1 = c(7,9,6),
                  p2 = c(10, 13, 11),
                  p3 = c(3.6, 4.833, 3.167),
                  p4 = c(6.667, 9.167, 6.6))
colnames(rtt) <- c("Parameters",
                   "$\\mathrm{CMPS_{max}}$ KNM's max",
                   "$\\mathrm{CMPS_{max}}$ KM's min",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KNM's max",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KM's min")

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
library(bulletxtrctr)
library(knitr)
library(kableExtra)

@@

3. **determine the congruent registration position:** \wj{A candidate position "receives" `votes` from basis segments that identify it or a close position within a tolerance zone of `Tx` as a candidate position in step 2. Votes for all candidate positions} are tallied and the position with the highest number of votes gets chosen as the *congruent registration position*. Basis segments with a congruent registration position are called "congruent matching profile segments" (CMPS). The total number of CMPS is the CMPS score of the comparison. \hh{XXX what happens when there are ties after the vote?}

@@

* `get_ccp(ccr.list, Tx = 25)` tries to identify the "consistent correlation peak" (defined in algorithm step 2) under the framework of multi segment lengths strategy. `ccr.list` is described in the previous bullet point, and `Tx` determines the size of a tolerance zone used in identifying the consistent correlation peak. \hh{XXX what happens when there is no consistent correlation peak?} `get_ccp(...)` returns `NULL` if there is no consistent correlation peak.

@@

The signatures of Land 4 of Bullet 1 and Land 5 of Bullet 2 are stored in objects `sigs2` and `sigs1`, respectively. This comparison is one of the KM comparisons. We compute the CMPS score using two versions of the CMPS algorithm:

@@

```{r hamby252-data, eval=TRUE, echo=FALSE}
# for hamby 252
N <- 3
CMPS_hamby252_results <- list()
CMPS_hamby252_results$span1 <- list(0.75, 0.25, 0.25)
CMPS_hamby252_results$signame <-
  list("sigs75", "sigs25", "sigs25_1062")
CMPS_hamby252_results$npeaks.set <-
  list(c(5, 3, 1),
       c(5, 3, 1),
       c(10, 6, 2))
CMPS_hamby252_results$seg_length <- list(50, 50, 50)
CMPS_hamby252_results$Tx <- list(25, 25, 25)
CMPS_hamby252_results$titlee <- list()
CMPS_hamby252_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby252_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", span1=",
      CMPS_hamby252_results$span1[[i]],
      ", span2=0.03, len=",
      CMPS_hamby252_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby252_results$Tx[[i]]
    )
  CMPS_hamby252_results$filename[[i]] <- 
    paste(
      "hamby252",
      CMPS_hamby252_results$span1[[i]]*100,
      paste(CMPS_hamby252_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby252_results$seg_length[[i]],
      CMPS_hamby252_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby252_results$cmps.table <- list()
CMPS_hamby252_results$plot <- list()

for(i in 1:N){
  CMPS_hamby252_results$cmps.table[[i]] <- 
  read.csv(file = paste("./data/", 
                        CMPS_hamby252_results$filename[[i]], 
                        ".csv",
                        sep = ""))
}

# generate plots
for (i in 1:3) {
  hamby252.cmps <- CMPS_hamby252_results$cmps.table[[i]]

  hamby252.plot.list <- list()
  titlee <- CMPS_hamby252_results$titlee[[i]]

  hamby252.plot.list[[1]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.max.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      fill = "Comparison Type",
      x = expression(CMPS[max]),
      title = expression(paste("Hamby252 - ", CMPS[max], " Distribution")),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 27, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())

  hamby252.plot.list[[2]] <- hamby252.cmps %>% ggplot() +
    geom_histogram(aes(x = cmps.maxbar.m,
                       fill = as.factor(type_truth)), binwidth = 1) +
    labs(
      x = expression(bar(CMPS)[max]),
      fill = "Comparison Type",
      title = expression(paste(
        "Hamby252 - ", bar(CMPS)[max], " Distribution"
      )),
      subtitle = titlee
    ) +
    scale_x_continuous(breaks = seq(0, 24, 1)) +
    theme_bw() +
    theme(panel.grid.minor = element_blank())

  plot <- ggarrange(plotlist = hamby252.plot.list,
                    nrow = 1,
                    ncol = 2)
  CMPS_hamby252_results$plot[[i]] <- plot
}

```

```{r hamby44-data, eval=TRUE, echo=FALSE}
N <- 3
CMPS_hamby44_results <- list()
CMPS_hamby44_results$span1 <- list(0.25, 0.25, 0.75)
CMPS_hamby44_results$signame <-
  list("sigs25_531", "sigs25_1061", "sigs75_531")
CMPS_hamby44_results$npeaks.set <-
  list(c(5, 3, 1),
       c(10, 6, 1),
       c(5, 3, 1))
CMPS_hamby44_results$seg_length <- as.list(rep(61, 3))
CMPS_hamby44_results$Tx <- as.list(rep(30, 3))
CMPS_hamby44_results$titlee <- list()
CMPS_hamby44_results$filename <- list()
for (i in 1:N) {
  CMPS_hamby44_results$titlee[[i]] <-
    paste0(
      "npeaks.set=c(",
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = ","),
      ")",
      ", span1=",
      CMPS_hamby44_results$span1[[i]],
      ", span2=0.03, len=",
      CMPS_hamby44_results$seg_length[[i]],
      ", Tx=",
      CMPS_hamby44_results$Tx[[i]]
    )
  CMPS_hamby44_results$filename[[i]] <- 
    paste(
      "hamby44",
      CMPS_hamby44_results$span1[[i]]*100,
      paste(CMPS_hamby44_results$npeaks.set[[i]], collapse = "-"),
      CMPS_hamby44_results$seg_length[[i]],
      CMPS_hamby44_results$Tx[[i]],
      sep = "_"
    )
}
CMPS_hamby44_results$cmps.table <- list()
CMPS_hamby44_results$plot <- list()
```

@@

In this example, $\mathrm{CMPS_{max}}$, the highest CMPS score, is 17, and $\mathrm{\overline{CMPS}_{max}}$, the highest possible mean CMPS score of comparisons in the same phase, is $(3+17+14+10+15+16)/6=12.5$.

@@

```{r, echo=FALSE, warning=FALSE}
# rtt <- readRDS("summary_table.rds")
# colnames(rtt) <- 
rtt <- data.frame(parameters = c("span1=0.25, (5, 3, 1)", "span1=0.25, (10, 6, 1)", "span1=0.75, (5, 3, 1)"),
                  p1 = c(7,9,6),
                  p2 = c(10, 13, 11),
                  p3 = c(3.6, 4.833, 3.167),
                  p4 = c(6.667, 9.167, 6.6))
colnames(rtt) <- c("Parameters",
                   "$\\mathrm{CMPS_{max}}$ KNM's max",
                   "$\\mathrm{CMPS_{max}}$ KM's min",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KNM's max",
                   "$\\mathrm{\\overline{CMPS}_{max}}$ KM's min")
